Fast and exact visibility on digitized shapes and application to saliency-aware normal estimation

### 2025-04-25 DGMM2025/paper.tex
tabular improvement
Working on 1

    To increase efficiency, several authors have proposed to cast the
    visibility problem into a digital space, generally $\Z^3$ or its
    decomposition into cubical cells. Visibility is simplified as the
    problem if there is a connected digital straight line joining a
    pair of cells without occluding cell(s) (e.g.\ Soille~\cite{soille:1994-prl}
    uses Bresenham line segments). Coeurjolly \emph{et al.}~\cite{coeurjolly:2004-prl}
    uses a non-symmetric visibility definition and preimage computations
    to speed up that kind of algorithms. Chica~\cite{chica:2008-spm}
    uses a half-cell erosion of the complementary space and Bresenham
    straight lines to decide visibility.

@@

    We focus on solving efficiently this problem because it is crucial
    in computing geometric normals along digital surfaces that are
    aware of salient features. Indeed, 2d discrete tangent estimators
    based on tangency or discrete line segments have proven to be both
    sensitive to sharp corners while multigrid convergent along the
    boundary of digitized smooth shapes~\cite{feschet:1999-dgci,lachaud:2007-ivc,nguyen:2011-pr}.
    Many ad hoc methods have been proposed for tangent plane or normal
    estimation on 3d digital surfaces
    (e.g. \cite{fourey:2009-cg,charrier:2011-iwcia,Cuel:2014-dgci,Lachaud:2017-lnm,mareche:2024-ispr}).
    If two of these methods~\cite{Cuel:2014-dgci,Lachaud:2017-lnm}
    have been shown to be multigrid convergent on digitization of
    shapes with smooth boundary, their formulation intrinsically
    smoothes features, e.g.\ in contrast with the method of
    Mar{ê}ch{é} \emph{et al.}~\cite{mareche:2024-ispr}, whose
    convergence is not established. We propose here a normal estimator
    defined as the most orthogonal vector to the cone of visibility at
    the point of interest. Visibility is high in smooth regions and
    the induced visibility vectors are good approximations of tangent
    vectors. On the contrary the visibility is blocked at sharp
    features, so the normal vector is not influenced by the
    geometry on the other side of the features. We demonstrate the
    superiority of this normal estimator for estimating curvatures
    along digital surfaces with sharp features.

    Note that our algorithm for computing visibility is in fact much
    more general. The same algorithm could be used to recognize if any
    pattern of lattice points is present on a shape. It is similar to
    an erosion algorithm where the structuring element is the chosen
    pattern, and could also be used for morphological operations~\cite{soille1999morphological}.
    The paper outline is as follows. In Section~2, we recall some
    definitions and present tangency of chords and its equivalence
    to visibility. Section~3 describes the specific data structure
    for encoding set of grid cells, and the main algorithm that
    exploits this encoding to compute visibility exactly. Section~4
    presents experimental results on visibility, its usage as a
    discrete normal estimator, and how it improves curvature estimates
    on digital surfaces. Finally, Section 5 concludes and gives some
    perspectives to this work, notably other usages of this algorithm
    in the field.

@@

    More precisely, we choose a Gaussian function
    $\Kernel{x}\coloneqq e^{-\frac{x^2}{2\sigma^2}}$ as a weight kernel. Let
    $V_p$ be the set of visible points from the point $p$. Let $c_p$
    be the weighted centroid of the visible points around $p$,
    i.e. $c_p \coloneqq \frac{\sum_{q \in V_p} \Kernel{\|q-p\|}q}{\sum_{q \in
    V_p} \Kernel{\|q-p\|}}$. We form the weighted covariance matrix
    $\mathcal{V}_p$ of the points $V_p$ as:
    \begin{equation}
        \mathcal{V}_p = \sum_{q \in V_p} \Kernel{\|q-p\|}(q - c_p)(q - c_p)^T.
    \end{equation}
    The \emph{visibility normal} $\vec{n}(p)$ of point $p$ at scale $\sigma$ is defined
    as the first eigenvector of the covariance matrix $\mathcal{V}_p$
    of the visible points, corresponding to its smallest
    eigenvalue. Its orientation is chosen so as to point in the same
    direction as the average of the trivial normals to the surfels
    touching the point $p$. Figure~\ref{fig:normals-estimation} shows an example
    of our normals compared to normals computed with the integral invariant (II) estimator~\cite{Lachaud:2017-lnm}.

    \begin{figure}
        \centering
        \begin{tabular}{|c||c|c|}
            \hline
            II normals &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-cubes-II} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-smooth-II} \\
            \hline
            Visibility normals &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-cubes-NV} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-smooth-NV} \\
            \hline
            II normals &
            \includegraphics[width=0.4\textwidth]{pictures/octaflower-normal-estimation-cubes-II} &
            \includegraphics[width=0.4\textwidth]{pictures/octaflower-normal-estimation-smooth-II} \\
            \hline
            Visibility normals &
            \includegraphics[width=0.4\textwidth]{pictures/octaflower-normal-estimation-cubes-NV} &
            \includegraphics[width=0.4\textwidth]{pictures/octaflower-normal-estimation-smooth-NV} \\
            \hline
        \end{tabular}
        \caption{Examples of normals computed on a chinese dragon, then on an
        octaflower as a color map. First set displays II normals, second set
        uses the visibility algorithm to compute them. The normals are colored
        according to their orientation as $0.5 \cdot (n_i + \mathbbm{1})$. The
        right images are the smooth version of the left images. ``Integral
        Invariant normals'' are computed using a radius of $r=3$, while
        our normal estimator are computed using a deviation of $\sigma=3$, and so
        a radius of $r=2\sigma=6$}
        \label{fig:normals-estimation}
    \end{figure}

>>>>

    To increase efficiency, several authors have casted the
    visibility problem into a digital space, generally $\Z^3$ or its
    decomposition into cubical cells. Visibility is simplified as the
    question whether there is a connected digital straight line
    joining a pair of cells without occluding cell(s)
    (e.g. Soille~\cite{soille:1994-prl} uses Bresenham
    segments). Coeurjolly \emph{et al.}~\cite{coeurjolly:2004-prl}
    uses a non-symmetric visibility definition and preimage
    computations to speed up these
    algorithms. Chica~\cite{chica:2008-spm} uses a half-cell erosion
    of the complementary space and Bresenham lines to decide
    visibility.

@@

    We focus on solving efficiently this problem because it is crucial
    in computing geometric normals along digital surfaces that are
    aware of salient features. Indeed, 2d discrete tangent estimators
    based on tangency or discrete line segments have proven to be both
    sensitive to sharp corners and multigrid convergent on the
    boundary of digitized smooth shapes~\cite{feschet:1999-dgci,lachaud:2007-ivc,nguyen:2011-pr}.
    Many ad hoc methods have been proposed for tangent plane or normal
    estimation on 3d digital surfaces
    (e.g. \cite{fourey:2009-cg,charrier:2011-iwcia,Cuel:2014-dgci,Lachaud:2017-lnm,mareche:2024-ispr}).
    If two of these methods~\cite{Cuel:2014-dgci,Lachaud:2017-lnm}
    have been shown to be multigrid convergent on digitization of
    shapes with smooth boundary, their formulation intrinsically
    smoothes features, e.g.\ in contrast with the method of
    Mar{ê}ch{é} \emph{et al.}~\cite{mareche:2024-ispr}, whose
    convergence is not established. We propose here a normal estimator
    defined as the most orthogonal vector to the cone of visibility at
    the point of interest. Visibility is high in smooth regions and
    the induced visibility vectors are good approximations of tangent
    vectors. On the contrary the visibility is blocked at sharp
    features, so the normal vector is not influenced by the
    geometry on the other side of the features. We demonstrate the
    superiority of this normal estimator for estimating curvatures
    along digital surfaces with sharp features.

    Note that our algorithm for computing visibility is in fact much
    more general. The same algorithm could be used to recognize if any
    pattern of lattice points is present on a shape. It is similar to
    an erosion algorithm where the structuring element is the chosen
    pattern, and could also be used for morphological operations~\cite{soille1999morphological}.
    The paper outline is as follows. In Section~2, we recall some
    definitions and present tangency of chords and its equivalence
    to visibility. Section~3 describes the specific data structure
    for encoding set of grid cells, and the main algorithm that
    exploits this encoding to compute visibility exactly. Section~4
    presents experimental results on visibility, its usage as a
    discrete normal estimator, and how it improves curvature estimates
    on digital surfaces. Finally, Section 5 concludes and gives some
    perspectives to this work.
    %% , notably other usages of this algorithm in the field.

@@

    More precisely, we choose a Gaussian function
    $\Kernel{x}\coloneqq e^{-\frac{x^2}{2\sigma^2}}$ as a weight kernel. Let
    $V_p$ be the set of visible points from the point $p$. Let $c_p$
    be the weighted centroid of the visible points around $p$,
    i.e. $c_p \coloneqq \frac{\sum_{q \in V_p} \Kernel{\|q-p\|}q}{\sum_{q \in
    V_p} \Kernel{\|q-p\|}}$. We form the weighted covariance matrix
    $\mathcal{V}_p$ of the points $V_p$ as:
    \begin{equation}
        \mathcal{V}_p = \sum_{q \in V_p} \Kernel{\|q-p\|}(q - c_p)(q - c_p)^T.
    \end{equation}
    The \emph{visibility normal} $\vec{n}(p)$ of point $p$ at scale $\sigma$ is defined
    as the first eigenvector of the covariance matrix $\mathcal{V}_p$
    of the visible points, corresponding to its smallest
    eigenvalue. Its orientation is chosen so as to point in the same
    direction as the average of the trivial normals to the surfels
    touching the point $p$.

    \paragraph{Experimental validation.} [TO COMPLETE]
    Figure~\ref{fig:normals-estimation} shows an example
    of our normals compared to normals computed with the integral invariant (II) estimator~\cite{Lachaud:2017-lnm}.

    \begin{figure}
        \centering
        \begin{tabular}{|c||c|c|}
            \hline
            Normals & Cubes & Smooth \\
            \hline
            \hline
            \raisebox{18mm}{II} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-cubes-II} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-smooth-II} \\
            \hline
            \raisebox{18mm}{Ours} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-cubes-NV} &
            \includegraphics[width=0.4\textwidth]{pictures/chinese-dragon-normal-estimation-smooth-NV} \\
            \hline
            \raisebox{18mm}{II} &
            \includegraphics[width=0.4\textwidth]{pictures/octaflower-normal-estimation-cubes-II} &
            \includegraphics[width=0.4\textwidth]{pictures/octaflower-normal-estimation-smooth-II} \\
            \hline
            \raisebox{18mm}{Ours} &
            \includegraphics[width=0.4\textwidth]{pictures/octaflower-normal-estimation-cubes-NV} &
            \includegraphics[width=0.4\textwidth]{pictures/octaflower-normal-estimation-smooth-NV} \\
            \hline
        \end{tabular}
        \caption{Examples of normals computed on a chinese dragon, then on an
        octaflower as a color map. First set displays II normals, second set
        uses the visibility algorithm to compute them. The normals are colored
        according to their orientation as $0.5 \cdot (n_i + \mathbbm{1})$. The
        right images are the smooth version of the left images. ``Integral
        Invariant normals'' are computed using a radius of $r=3$, while
        our normal estimator are computed using a deviation of $\sigma=3$, and so
        a radius of $r=2\sigma=6$}
        \label{fig:normals-estimation}
    \end{figure}
